---
title: "Lab08"
author: "Nathan Holterhoff, Taylor Kelly, Anthony Liu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 1

## Tournament Results Summary

| club| opponent |result|
|----------|-----|-------|
Palmeiras| Atlético Mineiro|1:0|
Palmeiras |Internacional |2:1|
Atlético Mineiro| Palmeiras |0:0|
Atlético Mineiro |Internacional| 1:1|
Internacional |Palmeiras |2:3|
Internacional| Atlético Mineiro| 1:2|

## 1.1 

### 1.1.1

This data is in long form. This is because each row is a single observation and there are identifiers that are repeated in rows.

### 1.1.2

| Club              | vs Palmeiras | vs Atlético Mineiro  | vs Internacional |
|-------------------|--------------|----------------------|------------------|
| Palmeiras  |-|1:0,0:0|2:1,3:2|
| Atlético Mineiro |0:0,1:0|-|1:1, 2:1|
| Internacional | 2:1, 1:2|1:2, 2:1|-|


## 1.2

|Name |NASDAQ |established |revenue ($B)| net income| assets|
|-|-|-|-|
|Amazon |AMZN| 1994.00| 574.00| 30.40| 528.00|
|Google |GOOG| 1998.00| 307.00| 79.80| 402.00|


### 1.2.1

This data is wide format since there are no repeated identifiers and there are multiple variables per row.

### 1.2.2

| Company | Variable        | Value |
|---------|-----------------|-------|
| Amazon  | NASDAQ          | AMZN  |
| Amazon  | Established     | 1994  |
| Amazon  | Revenue (\$B)   | 574   |
| Amazon  | Net Income (\$B)| 30.4  |
| Amazon  | Assets (\$B)    | 528   |
| Google  | NASDAQ          | GOOG  |
| Google  | Established     | 1998  |
| Google  | Revenue (\$B)   | 307   |
| Google  | Net Income (\$B)| 79.8  |
| Google  | Assets (\$B)    | 402   |

# 2

## 2.1

```{r}
path <- "UAH-lower-troposphere-wide_2023-11-07.csv"
surfaceData <- read_delim(path) %>%
  select(year, month, globe, nopol, sopol)

surfaceData
```
The data above is good.

## 2.2

This data is wide format since there are no repeated identifiers and there are multiple variables per row.

## 2.3

```{r}
longSurfaceData <- surfaceData %>%
  pivot_longer(
    cols = c(globe, nopol, sopol),
    names_to = "region",
    values_to = "temperature"
  )

longSurfaceData
```

## 2.4

```{r}
wideSurfaceData <- longSurfaceData %>%
  pivot_wider(
    names_from = region,        # Use 'region' as the source of the new column names
    values_from = temperature   # Fill the columns with 'temperature' values
  )

wideSurfaceData
```

The result above looks like the original data.